---
title: "A Guide to Data Science and Machine Learning"
author: "Hannah Rozenbaum"
output: pdf_document
---

```{r setup, include=FALSE}
install.packages(tidytext)
library(knitr)
opts_chunk_set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(DBI)
library(RSQLite)
library(tidytext)
```

# 1. Sampling Strategies & Corpus Formation

## Core Rules

- **Avoid Convenience Sampling:** Never default to the first files listed. Intentionally sample across key metadata fields (regions, demographics, date, etc) for representation .
- **Start with MVP Sizing:** Build pipelines on a small corpus (5–10 files) to run working code efficiently before scaling to a massive corpus.
- **Document Sampling Criteria:** State explicitly in documentation how files were chosen to ensure reproducibility and acknowledge selection limitations.

## **Strategy**

**Stratified Random Sampling / Balanced Representative Sampling**

1.  **Identify Key Metadata Dimensions:** Examine the collection's metadata tags, such as Date or Location. You may need to extract them manually if there is no convenient tagging system.

2.  **Sample Across Stratas:** If the archive spans 1950–2010 across 10 states, intentionally pick 1–2 transcripts from each decade or region rather than clustering in one area.

3.  **Document Your Sampling Rules:** In the README explicitly state your criteria: "To minimize selection bias, 10 transcripts were selected via stratified sampling across three distinct geographic regions (South, Midwest, Urban Northeast) and across two time periods (1950–1970 vs. post-1970)."

## Reference Code: Stratified Sample Verification

```{r sampling-check, eval=FALSE}
# Example verification check on raw metadata prior to processing
metadata_sample <- tibble(
  interview_id = 1:10,
  decade = c("1960s", "1960s", "1970s", "1970s", "1980s", "1980s", "1990s", "1990s", "2000s", "2000s"),
  region = rep(c("South", "Northeast", "Midwest", "West", "South"), 2)
)

# Confirm sample balance across dimensions
metadata_sample %>% 
  count(decade, region)

```

------------------------------------------------------------------------

# 2. Ingestion & Storage Standards

## Core Rules

- **Keep `data/raw/` Immutable:** Raw files should never be directly edited. Treat them as read-only.
- **Relational Storage First:** Ingest data into a structured database (e.g., SQLite).
- **Enforce Consistent Keying:** Assign unique ID(s) across all tables for clean join operations.

## Reference Code: SQLite Database Ingestion

```{r sqlite-ingestion, eval=FALSE}
# 1. Connect to local SQLite database
con <- dbConnect(RSQLite::SQLite(), "data/processed/testimonies.db")

# 2. Define schema and create table
dbExecute(con, "
  CREATE TABLE IF NOT EXISTS raw_transcripts (
    interview_id TEXT PRIMARY KEY,
    speaker TEXT,
    year INTEGER,
    region TEXT,
    transcript_text TEXT
  )
")

# 3. Disconnect when finished
dbDisconnect(con)

```

------------------------------------------------------------------------

# 3. Text Preprocessing & Cleaning Rules

## Core Rules

- **Regex Line Cleaning:** Strip out structural artifact patterns like page headers, page numbers, timestamp markers, etc.
- **Tokenize before Filtering:** Convert text to lower-case unigrams/bigrams prior to joining with stop-word lexicons.
- **Domain-Specific Stop Words:** Stopwords differ based on the type of documents you are analyzing (news article, spoken interview, movie script, etc). Build a custom exclusion list.

## Reference Code: Tidytext Cleaning Pipeline

```{r text-cleaning, eval=FALSE}
# Custom domain stop-words
custom_stops <- tibble(word = c("interviewer", "speaker", "transcript", "yeah", "um", "uh"))

# Standard cleaning pipeline
clean_tokens <- raw_data %>%
  # Remove timestamps and bracketed notes
  mutate(clean_text = str_remove_all(transcript_text, "\\[.*?\\]|\\(.*?\\)|\\d+:\\d+:\\d+")) %>%
  # Tokenize
  unnest_tokens(word, clean_text) %>%
  # Remove standard and custom stop-words
  anti_join(stop_words, by = "word") %>%
  anti_join(custom_stops, by = "word") %>%
  # Strip numbers and punctuation leftovers
  filter(!str_detect(word, "^[0-9]+$"))

```

------------------------------------------------------------------------

# 4. NLP Modeling & Evaluation Techniques

## Core Rules

- **TF-IDF for Term Importance:** Use Term Frequency-Inverse Document Frequency (TF-IDF) to highlight domain-distinctive words rather than high-frequency generic terms.
- **Topic Modeling (LDA):** Use Latent Dirichlet Allocation to discover latent themes across document groupings.
- **Track Sentiment Trajectories:** Evaluate narrative arcs across document progression (e.g., dividing long transcripts into 10% index chunks rather than taking a single average score).

## Reference Code: TF-IDF Calculation

```{r tf-idf, eval=FALSE}
# Compute distinctive terms per interview
interview_tfidf <- clean_tokens %>%
  count(interview_id, word, sort = TRUE) %>%
  bind_tf_idf(word, interview_id, n) %>%
  arrange(desc(tf_idf))

```
